---
title: "Lake Kivu potato"
author: "Yanru Xing"
date: "`r format(Sys.time(), '%d %B, %Y')`"

output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    df_print: paged
---

# Load packages

```{r, fig.width=14, fig.height= 8, echo=FALSE}
library(igraph)
library(tidyverse)
library(viridis)
library(scales)
library(rgdal)
library(rworldmap)
library(sf)
library(tmap)
library(raster)
library(INA)
library(maptools)
library(maps)


```

# Cropland connectivity

```{r, fig.width=10, fig.height= 10, echo=FALSE}
PotatoAFRICA <- raster('spam2017V1r1_SSA_gr_H_POTA_A.tif') # load cropland data
plot(PotatoAFRICA)
palette1 <- rev(viridis_pal(option = "B")(1000))

colid<-c(1:100)*10
colid <- c(1, colid )
palette1<-palette1[colid]
#----------------------------------------------

colPalletNew <- c( "#F4E156FF", "#F6D746FF", "#F8CD37FF", "#FAC329FF", "#FBB91EFF", "#FCAF13FF", 
                   "#FCA50BFF", "#FB9C06FF", "#FA9207FF", "#F8890CFF", "#F68013FF", "#F37819FF",
                   "#F06F20FF", "#EC6727FF", "#E85F2EFF", "#E25834FF", "#DD5139FF", "#D74B3FFF",
                   "#D04545FF", "#CA404AFF", "#C33B4FFF", "#BC3754FF", "#B43359FF", "#AC305EFF",
                   "#A42C60FF", "#9B2964FF", "#932667FF", "#922568FF", "#902568FF", "#8F2469FF",
                   "#8D2369FF", "#8C2369FF", "#8A226AFF", "#88226AFF", "#87216BFF", "#85216BFF",
                   "#84206BFF", "#82206CFF", "#801F6CFF", "#7F1E6CFF", "#7D1E6DFF", "#7C1D6DFF",
                   "#7A1D6DFF", "#781C6DFF", "#771C6DFF", "#751B6EFF", "#741A6EFF", "#721A6EFF",
                   "#71196EFF", "#6E196EFF", "#6D186EFF", "#6B186EFF", "#6A176EFF", "#68166EFF",
                   "#66166EFF", "#65156EFF", "#63156EFF", "#61136EFF", "#60136EFF", "#5E126EFF",
                   "#5C126EFF", "#5B126EFF", "#59106EFF", "#58106EFF", "#560F6DFF", "#540F6DFF",
                   "#530E6DFF", "#510E6CFF", "#500D6CFF", "#4D0D6CFF", "#4C0C6BFF", "#4A0C6BFF",
                   "#490B6AFF", "#470B6AFF", "#450A69FF", "#440A68FF", "#420A68FF", "#400A67FF",
                   "#3E0966FF", "#3D0965FF", "#3B0964FF", "#390963FF", "#380962FF", "#360961FF",
                   "#340A5FFF", "#320A5EFF", "#310A5CFF", "#2F0A5BFF", "#2D0B59FF", "#2B0B57FF",
                   "#290B55FF", "#280B53FF", "#250C51FF", "#240C4EFF", "#230C4BFF", "#200C49FF",
                   "#1F0C47FF", "#1D0C44FF", "#1C0C42FF", "#1A0C40FF", "#190C3DFF", "#170C3BFF",
                   "#150B38FF", "#150B36FF", "#130A33FF", "#110A31FF", "#11092EFF", "#0F092CFF",
                   "#0D082AFF", "#0C0827FF", "#0B0725FF", "#0A0723FF", "#090620FF", "#08051EFF",
                   "#07051CFF", "#060419FF", "#050418FF", "#040315FF", "#040312FF", "#030210FF",
                   "#02020EFF", "#02020CFF", "#02010AFF", "#010108FF", "#010106FF", "#010005FF",
                   "#000004FF", "#000004FF", "#000004FF")

# Africa_ext <- extent(-19, 52, -38,22)
# 
# plot(
#   crop(PotatoAFRICA, Africa_ext) / 10000,
#   col = colPalletNew,
#   xaxt = 'n',
#   yaxt = 'n',
#   axes = F,
#   box = F,
#   main = paste('cropland harvested area: potato africa mapspam2017'),
#   cex.main = 1
#   )
# 
# 
# plot(countriesLow, add=TRUE)
# 
# PotatoAFRICA <- crop(PotatoAFRICA, Africa_ext) / 10000 # Change value as cropland harvested area fraction
```

# Crop cropland area of Burundi and Rwanda

```{r}


s.sf_Burundi <- st_read("bdi_admbnda_adm0_igebu_ocha_itos_20171103.shp")
potato_bdi <- crop(PotatoAFRICA, s.sf_Burundi)

s.sf_Rwanda <- st_read("rwa_adm0_2006_NISR_WGS1984_20181002.shp")
potato_rwa <- crop(PotatoAFRICA, s.sf_Rwanda)

potato_bdi_rwa <- merge(potato_bdi, potato_rwa)
plot(potato_bdi_rwa)
plot(countriesLow, add = TRUE)
map('lakes', add=TRUE, fill=TRUE, col='skyblue', boundary='black')

# 
# s.sf <- st_read("ken_admbnda_adm1_iebc_20191031.shp")
# CCRI <- crop(meanmapsraster, s.sf)
# County3 <- s.sf[c(32,34,47), ]
# County3 <- s.sf[c(26), ]

```

```{r}

potato_bdi_rwaValue <- getValues(potato_bdi_rwa)
potatoID <- which(potato_bdi_rwaValue > 0)

length(potatoID)

#-------------------------------------------------------------


cropValue <- potato_bdi_rwaValue

potatoAggID <- which(potato_bdi_rwaValue > 0)
cell_id <- potatoAggID

length(potatoAggID)
#writeRaster(potatoEthAgg, "potato_bdi_rwa_5min_cropland_density.tif")

#--------------------------------------

Xmin <- NULL
Xmax <- NULL
Ymin <- NULL
Ymax <- NULL

for (i in 1:length(potatoAggID)) {
  extCells <- extentFromCells(potato_bdi_rwa, potatoAggID[i])
  Xmin = c(Xmin, extCells[1])
  Xmax = c(Xmax, extCells[2])
  Ymin = c(Ymin, extCells[3])
  Ymax = c(Ymax, extCells[4])
}
#-------------------------------------------
Potato_crop_land <- cropValue[potatoAggID]

points_crop <- data.frame(Xmin, Ymax, Potato_crop_land)
#write.csv(points_crop, "Potato BDI RWA GIS and cropland 5min.csv" )


cropValue <- Potato_crop_land
```

# Distance matrix

```{r}
distancematr <-
  dist(points_crop[ , c(1:2)],
  method = 'euclidean',
  diag = T,
  upper = T)


#View(distancematr)

distance_matrix <- as.matrix(distancematr)
GIS_pepper_sub<-points_crop[ , c(1:2)]

lon <- GIS_pepper_sub$Xmin
lat <- GIS_pepper_sub$Ymax

```

```{r}
## 1.2 Customize crop and values of parameters
beta0<-0.5                                       ###
beta<-1                                          ###
beta1<-1.5                                       ###
gamma00<-0.05                                    ###
gamma0<-0.1                                      ###
gamma<-0.2                                       ###
gamma1<-0.3                                      ###
gamma2<-1                                        ###
crop<-'potato' 
cutoff1<- 0  #cropland density cutoff        ###
cutoff2 <- 0
cutoff3 <- 0
cutoffadja <- 0
cutoffadja1 <- 0# cutoff of adjancecy matrix     ###
cutoffadja2 <- 0
cutoffadja3 <- 0
```

#  Surveillance analysis: smartsurv analysis

smartsurv is another evaluating the value of nodes for surveillance in studying
the invasion of a bioentity as part of a smart surveillance strategy.

The stoch = TRUE is selected, then the entries in the matrix are taken as
probabilities and within a realization an adjacency matrix of 1s and 0s is
generated based on these probabilities.  The number of realizations for analysis
was nrealz = 10.

Note that it is assumed that the diagonal of the adjacency matrix is 1 (i.e.,
that once the bioentity reaches a location it stays there) – otherwise, the
stopping algorithm will not function correctly.

## smartsurv

Perforem smartsurv analysis for the observed Kenya local seed potato trade
network.


# CCRI calculated by Inverse power-law function 

```{r ,fig.width=11.75, fig.height=6.0, dpi=150}
#CCRI BY Inverse power-law function 
#This version is revised on 07/23/2020

CCRI_smarSurv_powerlaw_function <- function(beta, cutoffadja, distance_matrix, cropValue)   {
    ##############################################
    #### create adjacency matrix
    
    distancematr <- distance_matrix # pairwise distance matrix
  #---- end of code
    distancematrexp <- distancematr^(-beta) #use function C=AX^(-beta), here A=1, X=distancematr
    cropmatr <- cropValue # complete gravity model with crop data
    cropmatr1 <- matrix(cropmatr, , 1 )
    cropmatr2 <- matrix(cropmatr, 1, )

    cropmatrix <- cropmatr1 %*% cropmatr2
    cropmatrix <- as.matrix(cropmatrix)
    
    #-------------------------------------------
    # Rescale cropmatrix and distancematrexp to give them the same weight in the gravity model
    cropmatrix <- cropmatrix/max(cropmatrix)
    diag(distancematrexp) <- 0
    distancematrexp <- distancematrexp/ max(distancematrexp)
    #-------------------------------------------
    cropdistancematr <- distancematrexp * cropmatrix # adjacecy matrix
  
    is.matrix(distancematrexp)
    logicalmatr <- cropdistancematr > cutoffadja # adjacency matrix after threshold
    stan <- cropdistancematr * logicalmatr
    stan <- round(stan, 6) # use round() because betweenness() may have problem when do the calculation
    cropdistancematrix <- graph.adjacency(stan,mode=c("undirected"),diag=F,weighted=T)#create adjacency matrix
    #----------------------------------------------
    #------------------------ smartsurv ------------------------
    # 20211109 add smartsurv
    stan_smart <- stan
    diag(stan_smart) <- 1
    A1 <- smartsurv(adjmat=stan_smart, stoch=T, nrealz=5)
    A1mean <- A1$meanarr
    smartsurv_score_meanarr <- colMeans(A1mean)
    #range(smartsurv_score_meanarr) 
    
    ##############################################
   ## sum of nearest neighbors degree
   knnpref0<-graph.knn(cropdistancematrix,weights=NA)$knn
   knnpref0[is.na(knnpref0)]<-0
   degreematr<-degree(cropdistancematrix)
   knnpref<-knnpref0*degreematr
   if(max(knnpref)==0){knnprefp=0}else
       if(max(knnpref)>0){knnprefp=knnpref/max(knnpref)/6}

   ##############################################
   #### node degree, node strengh 
   ####
   nodestrength<-graph.strength(cropdistancematrix) 
   nodestrength[is.na(nodestrength)]<-0
     if(max(nodestrength)==0){nodestr=0}else
        if(max(nodestrength)>0){nodestr=nodestrength/max(nodestrength)/6}
   ##############################################
   #### betweenness centrality
   #### 
   between<-betweenness(cropdistancematrix)
   between[is.na(between)]<-0
    if(max(between)==0){betweenp=0}else
        if(max(between)>0){betweenp=between/max(between)/2}
   ##############################################
   #### eigenvector and eigenvalues
   #### 
   eigenvectorvalues<-evcent(cropdistancematrix)
   ev<-eigenvectorvalues$vector
    ev[is.na(ev)]<-0
    if(max(ev)==0){evp=0}else
        if(max(ev)!=0){evp=ev/max(ev)/6}
   ##############################################
   #### CCRI is a weighted mean of 4 network metric
   ####    
 index<-knnprefp+evp+betweenp+nodestr
   
#    indexpre<-cropRaster
# #   indexpre[]<- NaN
#    indexpre[CellNumber]<- index
#    indexv<-indexpre
#    return(indexv)
   
   return(list(index,smartsurv_score_meanarr))
   }

```

# CCRI calculated by negative exponential function 

```{r ,fig.width=11.75, fig.height=6.0, dpi=150}

CCRI_smarSurv_negExponential_function <-function(gamma,cutoffadja, distance_matrix, cropValue)   {
    ##############################################
    #### create adjacency matrix
    ####
    distancematr <- distance_matrix
   
    
    eulernumber <- exp(1)
    distancematrexponential <- eulernumber ^ (-gamma * distancematr)# exponential model
    cropmatr <- cropValue # complete gravity model with crop data
    cropmatr1 <- matrix(cropmatr,,1) # complete gravity model with crop data
    cropmatr2 <- matrix(cropmatr,1,)
    cropmatrix <- cropmatr1 %*% cropmatr2
    cropmatrix <- as.matrix(cropmatrix)
        #-------------------------------------------
    # Rescale cropmatrix and distancematrexp to give them the same weight in the gravity model
    cropmatrix <- cropmatrix/max(cropmatrix)
    diag(distancematrexponential) <- 0
    distancematrexponential <- distancematrexponential/ max(distancematrexponential)
    #-------------------------------------------
    
    cropdistancematr <- distancematrexponential * cropmatrix
    logicalmatr <- cropdistancematr > cutoffadja
    stan <- cropdistancematr * logicalmatr
    stan <- round(stan, 6) # use round() because betweenness() may have problem when do the calculation
    cropdistancematrix<-graph.adjacency(stan,mode=c("undirected"),diag=F,weighted=T)#create adjacency matrix
    #------------------------ smartsurv ------------------------
    # 20211109 add smartsurv
    stan_smart <- stan
    diag(stan_smart) <- 1
    A1 <- smartsurv(adjmat=stan_smart, stoch=T, nrealz=5)
    A1mean <- A1$meanarr
    smartsurv_score_meanarr <- colMeans(A1mean)
    #range(smartsurv_score_meanarr) 
    
    ##############################################
    #### create network for all the selected nodes
    ####
    #V(cropdistancematrix)$color=colororder
    V(cropdistancematrix)$label.cex=0.7
    edgeweight<-E(cropdistancematrix)$weight*4000
    E(cropdistancematrix)$color="red"
    #plot(cropdistancematrix,vertex.size=povalue*300,edge.arrow.size=0.2,edge.width=edgeweight,vertex.label=NA,main=paste(crop, sphere1, 'adjacency matrix threshold>',cutoffadja, ', beta=',beta)) # network with weighted node sizes
   # plot(cropdistancematrix,vertex.size=5,edge.arrow.size=0.2,edge.width=edgeweight,vertex.label=NA,main=paste(crop, sphere1, 'adjacency matrix threshold>',cutoffadja, ', beta=',beta)) # network with identical node size
   knnpref0<-graph.knn(cropdistancematrix,weights=NA)$knn
   knnpref0[is.na(knnpref0)]<-0
   degreematr<-degree(cropdistancematrix)
   knnpref<-knnpref0*degreematr
   if(max(knnpref)==0){knnprefp=0}else
       if(max(knnpref)>0){knnprefp=knnpref/max(knnpref)/6}

   ##############################################
   #### node degree, node strengh 
   ####
   nodestrength<-graph.strength(cropdistancematrix) 
   nodestrength[is.na(nodestrength)]<-0
     if(max(nodestrength)==0){nodestr=0}else
        if(max(nodestrength)>0){nodestr=nodestrength/max(nodestrength)/6}
   ##############################################
   #### betweenness centrality
   #### 
   between<-betweenness(cropdistancematrix)
   between[is.na(between)]<-0
    if(max(between)==0){betweenp=0}else
        if(max(between)>0){betweenp=between/max(between)/2}
   ##############################################
   #### eigenvector and eigenvalues
   #### 
   eigenvectorvalues<-evcent(cropdistancematrix)
   ev<-eigenvectorvalues$vector
    ev[is.na(ev)]<-0
    if(max(ev)==0){evp=0}else
        if(max(ev)!=0){evp=ev/max(ev)/6}
   ##############################################
   #### plot index layer
   ####    
 index<-knnprefp+evp+betweenp+nodestr
# #   indexpre[]<- NaN   
#    indexpre<-cropRaster
#    indexpre[CellNumber] <- index
#    indexv<-indexpre
#    return(indexv)
  return(list(index,smartsurv_score_meanarr))

   }

```


## sensitivity analysis CCRI BY Inverse power-law function and negative exponential 

```{r ,fig.width=11.75, fig.height=6.0, dpi=150}
index1 <- CCRI_smarSurv_powerlaw_function(beta = 0.5, cutoffadja, distance_matrix, cropValue )
index2 <- CCRI_smarSurv_powerlaw_function(beta = 1, cutoffadja, distance_matrix, cropValue)
index3 <- CCRI_smarSurv_powerlaw_function(beta = 1.5, cutoffadja, distance_matrix, cropValue)
```


```{r}
index4 <- CCRI_smarSurv_negExponential_function(gamma = 0.05, cutoffadja, distance_matrix, cropValue)
index5 <- CCRI_smarSurv_negExponential_function(gamma = 0.1, cutoffadja, distance_matrix, cropValue)
index6 <- CCRI_smarSurv_negExponential_function(gamma = 0.2, cutoffadja, distance_matrix, cropValue)
index7 <- CCRI_smarSurv_negExponential_function(gamma = 0.3, cutoffadja, distance_matrix, cropValue)
index8 <- CCRI_smarSurv_negExponential_function(gamma = 1, cutoffadja, distance_matrix, cropValue)
```

# Complete sensitivity analysis of CCRI 

```{r ,fig.width=9, fig.height=8, dpi=150}

mean_index <- (index1[[1]]+index2[[1]]+index3[[1]]+index4[[1]]+index5[[1]]+index6[[1]]+index7[[1]]+index8[[1]]) / 8

indexpre <- potato_bdi_rwa
indexpre[cell_id] <- mean_index
mean_index_raster <- indexpre
#writeRaster(mean_index_raster, "CCRI_potato_bdi_rwa_5min_euclidean_distance.tif")

```


# Complete sensitivity analysis of smartsurv score

```{r ,fig.width=9, fig.height=8, dpi=150}

smartsurvScore_index <- (index1[[2]] +index2[[2]] +index3[[2]] +index4[[2]] +index5[[2]] +index6[[2]] +index7[[2]] +index8[[2]]) / 8

indexpre_score <- potato_bdi_rwa

indexpre_score[cell_id] <- smartsurvScore_index
smartsurv_score_raster <- indexpre_score
#writeRaster(smartsurv_score_raster, "smartsurv_score_potato_bdi_rwa_5min_euclidean_distance.tif")

```




